### 2011-08-30 Text/paper.tex
Added results of the mass transfer. Still too large coefficient.

\documentclass{article}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{graphicx}
\usepackage{amssymb,amsmath}
\usepackage{bm}
\usepackage{color}
\usepackage{tabularx}
\newcommand{\beq}{\begin{equation}}
\newcommand{\feq}{\end{equation}}
\newcommand{\beqstar}{\begin{equation*}}
\newcommand{\feqstar}{\end{equation*}}                                     
\newcommand{\beqal}{\begin{equation}\begin{aligned}}
\newcommand{\feqal}{\end{aligned}\end{equation}}
\newcommand{\beqalstar}{\begin{equation*}\begin{aligned}}
\newcommand{\feqalstar}{\end{aligned}\end{equation*}}
\begin{document}
\section{Motivation}
The mass transfer is a very important characteristic for microchannels  design. It allows to
properly manufacture a microchannel with necessary properties to ensure that chemical
reactions are performed in the best manner. In
what follows the mass transfer associated with the bubble moving in the microchannel through liquid
medium is studied.

@@

There are different methods to estimate the mass transfer coefficient $k_L$. We first examine the
theoretical definitions of the mass transfer in case of the point sources.
\subsection{Point mass sources}
Let us look at the infinitisemal small domain of the volume $A \Delta x$ in a steady-state, not
moving and with the mass source. Then the concentration difference can be found as $\Delta C = C^* -
C(t)$, where $C*$ is the imposed concentration on the infinitesimal concentration source, $C(t)$ is
the time depending average concentration. Therefore one can write the time dependent ODE for the
average concentration in the domain:
\beq
\dot{m}= A \Delta x \frac{\mathrm{d}C}{\mathrm{d} t} = k_L P (C^{*}-C(t)), 
\feq
with the initial condition $C(0)=0$
The solution in time becomes as follows:
\beq
C(t)= C^{*}(1-\exp(-k_L a t )), 
\feq
where $k_L a$ is the volumetric mass transfer coefficient.

This simple picture is able to predict mass transfer happening in moving with the velocity $U$
liquid, see Fig. \ref{fig:moving:frame}.
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/mass_transfer.eps}
\caption{The mass transfer in the moving liquid. \label{fig:moving:frame}}
\end{figure}

The mass accumulated in the unit volume $V=A \Delta x$ can be calculated as the difference of 
mass fluxes entering and leaving domain $U \bigl(C(x+\Delta x)-C(x)\bigr)$. We assume that the
source is a point source with the effective perimeter P. The accumulated mass should be proportional
to the mass transfer coefficient:
\beq
U \bigl(C(x+\Delta x)-C(x)\bigr)=k_L P (C^{*}-C(x)), 
\feq 
giving the same equation but only in the coordinate domain:
\beq
C(x)= C^{*}(1-\exp(-k_L a \frac{x}{U} )).
\label{main:mass:transfer:expression} 
\feq
The expression (\ref{main:mass:transfer:expression}) is exactly used in the experiments by
\citet{bercic-mass}. One should be accurate with the definition of velocities because we have two
different phases. Usually, if we cannot distinguish phases one can take the velocity $U$ to be as
average velocity or $U=U_{sl}+U_{sg}$, where $U_{sl}$ and $U_{sg}$ are liquid and gas superficial
velocities correspondingly. However, the real picture is much more complicated, as two phases can
have different mixing patterns from film to slug and from slug to film. For example, one of the
most important characteristics to properly analytically estimate the mass transfer coefficient is
the mixing patterns. For example, one can assume that the slug concentration is uniformly
distributed over the slug, which usually overpredicts the mass transfer \cite{irandoust} since some
tracer concentration from film is mixed with the slug. Another issue is that if the slug is long
enough then the film saturates with the tracer concentration and its influence on the mass transfer
can be negligible \cite{vanbaten-circular}. Mixing patterns of the film and the liquid slugs are of
great importance for the analytical estimation of the mass transfer \cite{yue-mass}.

\subsection{Numerical calculation}
In the numerical simulation it is not possible to simulate really long domains with the infinite
number of the bubbles. Therefore, one needs to come up with the simple smaller domain calculations
of the mass transfer coefficient. The mixing patterns are extremely important for the analytical
estimations of the mass transfer coefficient. This is not the case for numerical calculations,
where one needs to present the adequate picture and the way to calculate the volumetric mass
transfer coefficient. We consider the following possibilities of conducting numerical simulations:
\begin{enumerate}
 \item The tracer picking up procedure can be introduced in the manner of the considerations above.
We impose the concentration at the entrance to be constant $C(0,y)=C_{inlet}$, at the outlet we
impose the outflow boundary condition $\frac{\partial C}{\partial x}(L,y)=0$. One can se the sketch
of the approach in Fig. \ref{fig:benchmark:jos}. Without loss of the generality one can impose
$C_{inlet}=0$. After that one can estimate the mass transfer coefficient through the mass flux
depending on the coordinate $x$. The mass flux is calculated through the following equation:
\beq
j(x)=\int{U(x) C(x,y) \mathrm{d}y}.
\feq
As it was indicated the mass flux is proportional to the mass transfer coefficient $j(x)=k_L \Delta
C$, where the difference in the concentration is taken as $\Delta C=C^{*}-C_{media}$, where $C^{*}$
is the saturation concentration. The question remains with the $C_{media}$ which can be treated
differently. According to the Fick's law $C_{media}$ should be the concentration of the media
examined. However, that imposes the assumption of the ideal mixing in the liquid slug. Another
possibility to take $C_{media}$ as the outflow concentration. A few considerations:
\begin{itemize}
 \item Note, that in the presence of the
complicated geometries and non-homogenous concentrations distributions around the bubble one can
use any definitions. Though the definition of the mass averaged $C_{media}$ is prefered.  
\item Note, that if the tracer volume is large then $C_{media}$ can be close to zero.
\item Note, that one cannnot use the definition of the mass transfer coefficient as far as it
varies through the bubble due to inhomogenoities of the flow. Therefore, the volumetric mass
transfer coefficient is introduced that takes into the account the bubble geometry.
\end{itemize}

\begin{figure}
\includegraphics[width=\textwidth]{Figures/benchmark_jos.eps}
\caption{The numerical mass transfer benchmark. \label{fig:benchmark:jos}}
\end{figure}
\item The periodic domain.
\end{enumerate}

>>>>

\documentclass{article}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{graphicx}
\usepackage{amssymb,amsmath}
\usepackage{bm}
\usepackage{color}
\usepackage{tabularx}

\newcommand{\beq}{\begin{equation}}
\newcommand{\feq}{\end{equation}}
\newcommand{\beqstar}{\begin{equation*}}
\newcommand{\feqstar}{\end{equation*}}                                     
\newcommand{\beqal}{\begin{equation}\begin{aligned}}
\newcommand{\feqal}{\end{aligned}\end{equation}}
\newcommand{\beqalstar}{\begin{equation*}\begin{aligned}}
\newcommand{\feqalstar}{\end{aligned}\end{equation*}}
\newcommand{\vol}{k_L\,a}

\begin{document}
\section{Motivation}
The mass transfer is a very important characteristic for microchannels  design. It allows to
properly manufacture a microchannel with necessary properties to ensure that chemical
reactions are performed in the best manner. In
what follows the mass transfer associated with the bubble moving in the microchannel through liquid
medium is studied.

@@

There are different methods to estimate the mass transfer coefficient $k_L$. We first examine the
theoretical definitions of the mass transfer in case of the point sources.
\subsection{Point mass sources}
In what follows we will present three approaches to calculate the mass points mass transfer
coefficients:
\begin{enumerate}
\item
Let us look at the infinitisemal small domain of the volume $A \Delta x$ in a steady-state, not
moving and with the mass source. Then the concentration difference can be found as $\Delta C = C^* -
C(t)$, where $C^*$ is the imposed concentration on the infinitesimal concentration source, $C(t)$ is
the time depending average concentration. Therefore one can write the time dependent ODE for the
average concentration in the domain:
\beq
\dot{m}= A \Delta x \frac{\mathrm{d}C}{\mathrm{d} t} = k_L P (C^{*}-C(t)), 
\feq
with the initial condition $C(0)=0$
The solution in time becomes as follows:
\beq
C(t)= C^{*}(1-\exp(-k_L a t )), 
\feq
where $k_L a$ is the volumetric mass transfer coefficient.

\item
The previous consideration is able to predict mass transfer happening in moving with the velocity
$U$ liquid, see Fig. \ref{fig:moving:frame}.
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/mass_transfer.eps}
\caption{The mass transfer in the moving liquid. \label{fig:moving:frame}}
\end{figure}

The mass accumulated in the unit volume $V=A \Delta x$ can be calculated as the difference of 
mass fluxes entering and leaving domain $U \bigl(C(x+\Delta x)-C(x)\bigr)$. We assume that the
source is a point source with the effective perimeter P. The accumulated mass should be proportional
to the mass transfer coefficient:
\beq
U \bigl(C(x+\Delta x)-C(x)\bigr)=k_L P \bigl(C^{*}-C(x)\bigr), 
\feq 
giving the same equation but only in the coordinate domain:
\beq
C(x)= C^{*}\Bigl(1-\exp\bigl(-k_L a \frac{x}{U} \bigr)\Bigr).
\label{main:mass:transfer:expression} 
\feq

\item If one transfers to the frame moving with the liquid velocity $U$, then the situation will be
the same as in the first case. However, one can connect the time and spatial location with the
velocity $U$ ($t=\frac{x}{U}$) to obtain the same equation as in the case 2.
\end{enumerate}

The expression (\ref{main:mass:transfer:expression}) was used in the experiments by
\citet{bercic-mass}. One should be accurate with the definition of velocities because we have two
different phases. Usually, one can take the velocity $U$ to be as
average velocity or $U=U_{gas}+U_{liq}$, where $U_{gas}$ and $U_{liq}$ are liquid and gas
superficial velocities correspondingly. In this case, the continuouty assumption is restored and
all considerations above are valid.

 However, if one wants to analytically estimate the mass transfer coefficients, the situation is
much more complicated because of the two phases presence. Phases can have different mixing patterns
from film to slug and from slug to film. The analytical approaches described above assume that the
contributions from film and bubble caps can be calculated separately. However one can see that such 
an assumption for bubble caps (slug well mixed and the concentration is uniformly distributed over
the slug) usually overpredicts the mass transfer \cite{irandoust}. This happens since some tracer
concentration from film is mixed with the slug and changes the overall concentration in the slug.
Another issue is that if the slug is long enough then the film saturates with the tracer
concentration and its influence on the mass transfer can be negligible \cite{vanbaten-circular}.
Mixing patterns of the film and the liquid slugs are of great importance for the analytical
estimation of the mass transfer \cite{yue-mass}.

\subsection{Numerical calculation}
In the numerical simulation it is not possible to simulate really long domains with the infinite
number of the bubbles. As we saw before there are two approaches towards it - either to simulate
the bubble train and then to measure concentration along the pipe, Eq.
\ref{main:mass:transfer:expression}, or to transfer to the reference frame moving with bulk
velocity $U$ and conduct the same measurements through Eq. \ref{main:mass:transfer:expression}.
However, both methods do require a tracking of moving bubbles which is complicated from the
numerical point of view.

Therefore, one needs to come up with the simple smaller domain calculations
of the mass transfer coefficient, which represent the continuous picture with infinite numbers of
the bubbles closely. In comparison with analytical estimations of the mass transfer coefficient,
where the mixing patterns are extremely important, the numerical simulations need only to mimic
the continous picture and provide the way to calculate the volumetric mass transfer coefficient.

We perform simulations in the frame moving with the bubble (velocity
$U_{bubble}$), that the bubble position stands steady. The bubble velocity $U_{bubble}$ is
different from the bulk velocity $U_{gas}+U_{liq}$. In this case we need to perform coordinate
change:
\beqal
&x(t)=U_{bubble}t\\
&C(x)=C^{*}\Bigl(1-\exp\bigl(-k_L\, a \frac{x}{U_{gas}+U_{liq}}\bigr)\Bigr)\\
&C(t)=C^{*}\Bigl(1-\exp\bigl(-k_L\, a\, t \frac{U_{bubble}}{U_{gas}+U_{liq}}\bigr)\Bigr)
\feqal
If the unit volume that we simulate is quite long one can obtain the values of the concentration
in the inlet and outlet as a function of time:
\beqal
&C(U_{bubble}\, t)=C^{*}\Bigl(1-\exp\bigr(-k_L\, a\, t
\frac{U_{bubble}}{U_{gas}+U_{liq}}\bigr)\Bigr)\\
&C(U_{bubble}\,t+L_{unit})=C^{*}\Bigl(1-\exp\bigl(-k_L a \frac{
U_{bubble}\,t+L_{unit}}{U_{gas}+U_{liq}}\bigr)\Bigr)\\
&=C^{*}\Bigl(1-\exp\bigl(-k_L\, a\, t \frac{U_{bubble}}{U_{gas}+U_{liq}}\bigr)\exp\bigl(-k_L a
\frac{L_{unit}}{U_{gas}+U_{liq}}\bigr)\Bigr)
\feqal
Therefore the mass transfer coefficient can be calculated as follows:
\beqal
&-k_L a \frac{L_{unit}}{U_{gas}+U_{liq}}=\ln\Bigl(\frac{C^{*}-C_{outlet}}{C^{*}-C_{inlet}}\Bigr)\\
&k_L a
=\frac{U_{gas}+U_{liq}}{L_{unit}}\ln\Bigl(\frac{C^{*}-C_{inlet}}{C^{*}-C_{outlet}}
\Bigr)
\feqal 
If inlet and outlet concentrations are small in comparison with the saturation concentration $C^{*}$
then the expression for the volumetric mass transfer coefficitent $k_L a$ can be simplified:
\beq
\vol=\frac{\Delta C}{C^{*}} \frac{U_{gas}+U_{liq}}{L_{unit}},
\feq
where $\Delta C=C_{outlet}-C_{inlet}$. If we go to the continuous formulation by changing $L_{unit}$
to be a continuous coordinate $x$, then the following can be obtained \cite{jos-mass}:
\beq
\Delta C = C^{*} \vol \frac{x}{U_{gas}+U_{liq}}
\label{small:conc:mass:estimation}
\feq

We consider the following possibilities of conducting numerical simulations:
\begin{enumerate}
 \item 
The tracer picking up procedure can be introduced
in the manner of the considerations above. We impose the concentration at the entrance to be
constant $C(0,y)=C_{inlet}$, at the outlet we impose the outflow boundary condition $\frac{\partial
C}{\partial x}(L,y)=0$. To be compliant with the considerations above one needs to impose as well
the outflow boundary conditions for the inlet as well. However, without loss of the generality one
can impose $C_{inlet}=0$, since the walls in the reference frame
moving with the bubble are moving from the inlet to the outlet. That means that no tracer
concentration leaves domain from the inlet side, which is true for the capillary numbers larger
than $0.6$.  However, this boundary condition can not be used in the case of small capillary
numbers, since there is a strong circulation in front of the bubble. One can see the sketch of the
approach in Fig. \ref{fig:benchmark:jos}. The longer the numerical domain length the better will be
results. One can estimate the volumetric mass transfer coefficient in this case through the equation
(\ref{small:conc:mass:estimation}) for small concentrations. Note a few considerations:
\begin{itemize}
\item  The considerations above were established for the homogenous velocity field. In the case of
the non-homogenous velocity field one needs to obtain a concentration from the mass flux:
\beq
C_{outlet}=\frac{\int{U(y) C(L_{unit},y) \mathrm{d}y}}{\int{U(y) \mathrm{d}y}}.
\feq
\item Note, that one cannnot use the definition of the mass transfer coefficient as far as it
varies through the bubble due to inhomogenoities of the flow. Therefore, the volumetric mass
transfer coefficient is introduced that takes into the account the bubble geometry.
\end{itemize}
\begin{figure}
\includegraphics[width=\textwidth]{Figures/benchmark_jos.eps}
\caption{The numerical mass transfer benchmark. \label{fig:benchmark:jos}}
\end{figure}

\item Another possibility is to use the periodic boundary conditions as the limiting case of the
continuous formulation. In this case the inlet and outlet concentrations are not much different
($C(x)=C(x+L_{unit})=C(x+\Delta x)\approx C_{media}$, see Fig. \ref{fig:moving:frame}), where
$C_{media}$ is the mass averaged over domain concentration. This formulation requires the periodic
boundary condition $C_{inlet}=C_{outlet}$. By the definition the mass transfer coefficient is
proportional to the gain of the mass in the system divided by the concentration difference:
\beq
j=\frac{\dot{m}}{P \Delta C}=\frac{\dot{m}}{P (C^{*}-C_{media})},
\feq
where mass difference in the domain can be calculated as:
\beq
\dot{m}=\frac{m_2-m_1}{t_2-t_1}=\frac{\int_{C(t_2)\mathrm{d}V}-\int{C(t_1)\mathrm{d}V}}{t_2-t_1}.
\feq
Therefore the assumptions  of this approach is that the difference between inlet and outlet is not
considerably large and the tracer is well mixed in the slug, as one can take the concentration
difference as $\Delta C = C^{*}-C_{media}$. The best case scenario is when $C_{media}$ is close to
zero, in this case we can be sure that the flux to the domain is calculated precisely and periodic
boundary conditions are fulfilled.
\end{enumerate}
